---
title: "Fundamental data type in R"
author: "Mike Yang"
date: "10 November 2018"
output:
  html_document: default
---
```{r message=FALSE}
library(tidyverse)
```

##Summary

> Although R has object-oriented features, methods are not part of the class definitions. Methods in R look just like normal functions, such as summary().

> Data type in R

* __vector__
    + roughly, there are two different kinds of vectors: atomic vectors and lists (generic vectors).
    + a single number (e.g., 3) is treated as a vector with only one element.
  
* __list__
    + Virtually every object you'll run across in R is really just a list.
  
    + Most objects that you'll deal with have the type "list".
  
    + If you want to actually "extract" an element from a list, you need to use the "[[]]" style of indexing.
  
    + If you have named elements of a list, instead of using the "[[]]" style of indexing mentioned previously, you can access them by appending $name instead. 

* __matrix__ 

* __data frame__
    + data frames are lists of data representing different kinds of information taken at the same times, places, etc. 
    + When you read in data from an external source, it will typically be in a data frame.

##Create data type in R

* The easiest way to directly create a multi-element vector is using the __c()__ (concatenate) function.

* You can create lists directly using the __list()__ function.

* Matrices can be created directly using __matrix()__.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#original article from 
#https://www.nealgroothuis.name/introduction-to-data-types-and-objects-in-r/
```


### creating a list
```{r}
df <- list(a = c(1,2), b = c(2,4), c = "mikey")
summary(df)
```

### access list element
```{r}
df[1]
class(df[1])
```

```{r}
df[[1]]
class(df[[1]])
```

```{r}
df$a
class(df$a)
```

```{r}
# convert list to tibble
df_table <- as.tibble(df)
```

### create a matrix
```{r}
m <- matrix(c(1,2,3,4), nrow = 2, ncol = 2)
```

### return first row and first column of the matrix
```{r}
m[1,1]
```

### return column of the matrix
```{r}
m[,2]
```

### return row of the matrix
```{r}
m[2,]
```